Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-09 alloscore_manuscript/supplement.rnw
some edits to last elr pr and a comment

A desirable feature of a scoring rule is that it is \emph{proper}, which roughly means that it rewards a forecaster for stating their true beliefs about the events being forecasted. We give a more formal definition of propriety in section \ref{sec:proper_def} before addressing the propriety of the allocation score in three stages:
\begin{itemize}
  \item Section \ref{sec:alloscore_proper} shows that the allocation score is proper when the forecaster's actual forecast distribution $F$ is available (as is any score derived using the decision theoretic procedures outlined in section 2 of the manuscript)
  \item Section \ref{sec:distfromq_alloscore_proper_prospective} shows that the allocation score is still proper if the forecast is represented as a member of a parametric family of distributions, so long as this representation and scoring procedure was specified prospectively by the hub. This includes the use of methods such as \verb`distfromq` for constructing forecast distributions as a special case.
  \item Section \ref{sec:distfromq_alloscore_improper_post_hoc} shows that \emph{post hoc} application of this scoring procedure to quantile forecasts is improper, and makes some suggestions for organized forecasting efforts who are considering using the allocation score.
\end{itemize}

@@

We recall the three-step decision theoretic procedure for deriving proper scoring rules that we outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
\end{enumerate}

By construction, this procedure yields a proper score since
\begin{align}
\Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
 &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
 &= \Ex_F [ S(G, Y)]. \nonumber
\end{align}
Here, step~\eqref{eqn:dt_proper_key} holds because the Bayes act $x^F$ was selected so that the expected loss $\Ex_F [ s(x^F, Y) ]$ for that action is less than or equal to the expected loss $\Ex_F [ s(x, Y) ]$ for any other action $x$ (and for the action $x^G$ in particular).

Because the alloscore was developed using this procedure, it is a proper score.

@@

Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss:
\begin{align*}
\theta^F &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
&= \text{argmin}_\theta \Ex_F [s_A(x^{F_\theta}, Y)]
\end{align*}
This leads the forecaster to choose values of the parameters $\theta$ such that the allocation $x^{F_\theta}$ is as close as possible to the allocation $x^F$ under the forecaster's true forecast distribution $F$.

>>>>

In order for a scoring rule to be guaranteed to yield logically consistent forecast evaluations, it must be \emph{proper}, which means,
in essence, that a forecaster cannot perceive any forecast at odds with their beliefs about the events being forecasted as having an optimal expected score.
We formalize this in terms of forecaster ``self-assessment'' in section \ref{sec:proper_def} and then address the propriety of the allocation score in three stages:
\begin{itemize}
  \item Section \ref{sec:alloscore_proper} shows that the allocation score is proper when the forecaster's actual forecast distribution $F$ is available (as is any score derived using the decision theoretic procedures outlined in section 2 of the manuscript)
  \item Section \ref{sec:distfromq_alloscore_proper_prospective} shows that the allocation score is still proper if the forecast is represented as a member of a parametric family of distributions, so long as this representation and scoring procedure was specified prospectively by the hub. This includes the use of methods such as \verb`distfromq` for constructing forecast distributions as a special case.
  \item Section \ref{sec:distfromq_alloscore_improper_post_hoc} shows that \emph{post hoc} application of this scoring procedure to quantile forecasts is improper, and makes some suggestions for organized forecasting efforts who are considering using the allocation score.
\end{itemize}

@@

We recall the three-step decision theoretic procedure for deriving proper scoring rules that we outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
  \item Define a \emph{scoring rule} which assigns the loss realized by the Bayes act $x^F$ against outcome $y$ as the score of $F$ at $y$:
  \begin{align}
    S(F, y) = s(x^F, y). \label{eqn:bayes_sr}
  \end{align}
\end{enumerate}

Such scoring rules, which we call \emph{Bayes scoring rules}, are proper by construction since
\begin{align}
\Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
 &= \min_{x} \Ex_F [ s(x, Y) ] \quad \text{ (by definition of $x^F$)} \\
 &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
 &= \Ex_F [ S(G, Y)]. \nonumber
\end{align}

The allocation scoring rule is Bayes and therefore proper.

We note that in the probabilistic forecasting literature (see e.g., \cite{gneiting2011making}, Theorem 3) what we have 
termed Bayes scoring rules typically appear via \eqref{eqn:bayes_sr} where $x^F$ is some given functional of $F$ which 
can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$.
Such a loss function is said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important
recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss 
function for $x^F$ which is consistent. Our orientation
is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
and use the Bayes act only as a bridge to a proper scoring rule.  Consistency is never in doubt.

@@

Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss:
\begin{align*}
\theta^F &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
&= \text{argmin}_\theta \Ex_F [s_A(x^{F_\theta}, Y)]
\end{align*}
This leads the forecaster to choose values of the parameters $\theta$ such that the allocation $x^{F_\theta}$ is as close as possible to the allocation $x^F$ under the forecaster's true forecast distribution $F$.
\apg{I think we need to say something about what we mean by ``close''.  The score divergence is what we are confronting here and I appreciate that 
others may not want to broach the topic.  But I think we have to say something about ``score expectation defined distance''. This also relates directly to the notation crunch below.}
